---
title: "QBio305 - Bioclimatic data for the Arabidopsis thaliana accessions in Spain and Sweden"
author: "Ingo Giebel"
date: today
lang: en-US
eol: lf
resource-path: 
  - .
  - ..
  - ../docs
  - ../results
bibliography: bibliography.bib
csl: apa
pdf-engine: lualatex
latex-makeindex: xindex
self-contained: true
self-contained-math: true
title-block-banner: true
page-layout: full
toc: true
toc-location: left
toc-expand: 3
number-sections: true
code-tools: true
code-link: true
code-fold: true
code-overflow: wrap
fig-format: retina
fig-align: left
fig-responsive: true
fig-width: 9.5
fig-height: 6.5
smooth-scroll: true
format:
  html:
    theme:
      light: flatly
      dark: darkly
    grid:
      sidebar-width: 450px
      body-width: 900px
    html-math-method: mathjax
knitr:
  opts_chunk:
    warning: false
    message: false
    dev: svglite
    fig.path: ../results/
execute:
  cache: true
  freeze: auto
editor: visual
---

# Project folder structure

@fig-proj_folder_struct shows the project folder structure required to execute this Quarto script (highlighted in purple). The project's base directory (= working directory) is automatically set when this project is opened via the RStudio project file `qbio305_student-work-a-thaliana.Rproj`.

The `data` subfolder contains the data files. The `docs` subfolder contains the LaTeX research report and the BibLaTex file `bibliography.bib` which includes all referenced bibliography entries (both for the research report and this script). The `results` subfolder contains all figures and tables created by this script. And, finally, the `scripts` subfolder contains this script, the produced HTML file, and the `apa.csl` configuration file required for creating APA (7th edition) conformant citations and bibliography entries.

```{r, engine = "tikz"}
#| code-summary: "Draw the project folder structure"
#| label: fig-proj_folder_struct
#| fig-cap: "Folder structure of this project"
#| out-width: 12.0cm
#| out-height: 12.0cm

\begin{tikzpicture}[thick, font=\sffamily\large]

  \draw (0, -1) -- +(1, 0) coordinate (c_1)
    ++(0, -3) -- +(1, 0) coordinate (c_2)
    ++(0, -2) -- +(1, 0) coordinate (c_3)
    ++(0, -1) -- +(1, 0) coordinate (c_4)
    ++(0, -4) -- +(1, 0) coordinate (c_5);

  \node[right, draw, rounded corners] at (-1, 0) {project base dir};

  \node[right] at (c_1) {data};
  \draw (c_1) ++(0, -1) -- +(0.5, 0) coordinate (c_1_1)
    ++(0, -1) -- +(0.5, 0) coordinate (c_1_2);
  \node[right] at (c_1_1) {geodata};
  \node[right] at (c_1_2) {accessions.tsv};
  \draw (c_1) ++(-0.2, 0) -- +(0, -2);

  \node[right] at (c_2) {docs};
  \draw (c_2) ++(0, -1) -- +(0.5, 0) coordinate (c_2_1);
  \node[right] at (c_2_1) {bibliography.bib};
  \draw (c_2) ++(-0.2, 0) -- +(0, -1);

  \node[right] at (c_3) {results};

  \node[right] at (c_4) {scripts};
  \draw (c_4) ++(0, -1) -- +(0.5, 0) coordinate (c_4_1)
    ++(0, -1) -- +(0.5, 0) coordinate (c_4_2)
    ++(0, -1) -- +(0.5, 0) coordinate (c_4_3);
  \node[right] at (c_4_1) {apa.csl};
  \node[right] at (c_4_2) {04\_accessions\_climate\_data.html};
  \node[right, purple] at (c_4_3) {04\_accessions\_climate\_data.qmd};
  \draw (c_4) ++(-0.2, 0) -- +(0, -3);

  \node[right] at (c_5) {student-work-arabidopsis-thaliana.Rproj};

  \draw (0, -0.4) -- (0, -12);

\end{tikzpicture}
```

# R Packages and options

## Load the required packages

```{r appendix = TRUE}
#| code-summary: "Load the required packages"
#| output: false

# For determining the project root
library(here)
# For transforming, representing, and plotting data
library(tidyverse)
library(ggrepel)
# For printing tables
library(gt)
# For retrieving and plotting the geographical and climatic data
library(geodata)
library(terra)
library(tidyterra)
library(raster)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)
library(ggspatial)
```

## Set scientific notation options and default plotting theme

```{r appendix = TRUE}
#| code-summary: "Set scientific notation options and default plotting theme"
#| code-fold: false
#| output: false

options(scipen = 1)
options(pillar.sigfig = 4)

ggplot2::theme_set(ggplot2::theme_classic())
```

# Data import

## Data file paths

```{r appendix = TRUE}
#| code-summary: "Determine the file paths of the data files relative to the
#| project's base directory"
#| output: false

dta_accessions_file_path <- here::here(
  "data",
  "accessions.tsv")

dta_geo_dir_path <- here::here(
  "data",
  "geodata")

dta_uvb_file_path <- here::here(
  "data",
  "geodata",
  "56459_UVB1_Annual_Mean_UV-B.asc")

dta_pet_06_file_path <- here::here(
  "data",
  "geodata",
  "CHELSA_pet_penman_06_1981-2010_V.2.1.tif")
```

## Read in the data

```{r appendix = TRUE}
#| code-summary: "Read in the data"
#| output: false
#| cache: false

# Read in the accessions data
dta_accessions <- readr::read_tsv(
  file = dta_accessions_file_path,
  col_types = "ffnn")

# Extract a list of the accession coordinates
acc_coords <- dta_accessions[, c("lon", "lat")]

# Read in the bioclimatic data. If the data is not already present at the
# specified path, it will be downloaded.
dta_bio_climate <- geodata::worldclim_global(
  var = "bio",
  res = 0.5,
  path = dta_geo_dir_path)

# Read in the annual mean UVB index data
# Source: https://www.ufz.de/gluv/index.php?en=32367
dta_uvb_idx <- raster::raster(dta_uvb_file_path)

# Read in the potential evapo-transpiration (PET) monthly data (June)
# averaged for 1981-2010
# Source: https://envicloud.wsl.ch/#/?prefix=chelsa%2Fchelsa_V2%2FGLOBAL%2F
dta_pet_06 <- raster::raster(dta_pet_06_file_path)
```

# Extract the bioclimatic data

@tbl-accessions_climate shows the climate data for the accessions.

```{r appendix = TRUE}
#| code-summary: "Extract the bioclimatic data for the accession coordinates"
#| cache: false
#| label: tbl-accessions_climate
#| tbl-cap: "climate data for the accessions"

# Extract the WorldClim bioclimatic data for the accession coordinates
dta_bio_climate_extr <- terra::extract(
  dta_bio_climate,
  acc_coords,
  method = "bilinear",
  df = TRUE
)

# Extract the annual mean UVB index data for the accession coordinates
dta_uvb_idx_extr <- raster::extract(dta_uvb_idx, acc_coords, df = TRUE)

# Extract the potential evapo-transpiration (PET) monthly data (June)
dta_pet_06_extr <- raster::extract(dta_pet_06, acc_coords, df = TRUE)

# Create a tibble of the accessions data and the extracted bioclimatic data
dta_accessions_climate_df <- dplyr::bind_cols(
  dta_accessions |>
    dplyr::rename(
      "Sample" = "sample",
      "Population" = "pop",
      "Latitude" = "lat",
      "Longitude" = "lon"
    ),
  dta_bio_climate_extr |>
    dplyr::select(-1) |>
    dplyr::rename(
      "Annual mean temperature" = "wc2.1_30s_bio_1",
      "Mean diurnal range" = "wc2.1_30s_bio_2",
      "Isothermality" = "wc2.1_30s_bio_3",
      "Temperature seasonality" = "wc2.1_30s_bio_4",
      "Max temperature of warmest month" = "wc2.1_30s_bio_5",
      "Min temperature of coldest month" = "wc2.1_30s_bio_6",
      "Temperature annual range" = "wc2.1_30s_bio_7",
      "Mean temperature of wettest quarter" = "wc2.1_30s_bio_8",
      "Mean temperature of driest quarter" = "wc2.1_30s_bio_9",
      "Mean temperature of warmest quarter" = "wc2.1_30s_bio_10",
      "Mean temperature of coldest quarter" = "wc2.1_30s_bio_11",
      "Annual precipitation" = "wc2.1_30s_bio_12",
      "Precipitation of wettest month" = "wc2.1_30s_bio_13",
      "Precipitation of driest month" = "wc2.1_30s_bio_14",
      "Precipitation seasonality" = "wc2.1_30s_bio_15",
      "Precipitation of wettest quarter" = "wc2.1_30s_bio_16",
      "Precipitation of driest quarter" = "wc2.1_30s_bio_17",
      "Precipitation of warmest quarter" = "wc2.1_30s_bio_18",
      "Precipitation of coldest quarter" = "wc2.1_30s_bio_19"
    ),
  dta_uvb_idx_extr |>
    dplyr::select(-1) |>
    dplyr::rename(
      "UVB index" = "X56459_UVB1_Annual_Mean_UV.B"
    ),
  dta_pet_06_extr |>
    dplyr::select(-1) |>
    dplyr::rename(
      "PET (June)" = "CHELSA_pet_penman_06_1981.2010_V.2.1"
    )
)

# Table output
tbl_accessions_climate <- dta_accessions_climate_df |>
  gt::gt() |>
  gt::opt_stylize() |>
  gt::tab_options(
    table.align = "left",
    table.font.size = "80%"
  ) |>
  gt::fmt_number(
    columns = c(
      `Annual mean temperature`,
      `Mean diurnal range`,
      `Isothermality`,
      `Max temperature of warmest month`,
      `Min temperature of coldest month`,
      `Temperature annual range`,
      `Mean temperature of wettest quarter`,
      `Mean temperature of driest quarter`,
      `Mean temperature of warmest quarter`,
      `Mean temperature of coldest quarter`
    ),
    decimals = 1
  ) |>
  gt::fmt_integer(
    columns = c(
      `Temperature seasonality`,
      `Annual precipitation`,
      `Precipitation of wettest month`,
      `Precipitation of driest month`,
      `Precipitation seasonality`,
      `Precipitation of wettest quarter`,
      `Precipitation of driest quarter`,
      `Precipitation of warmest quarter`,
      `Precipitation of coldest quarter`,
      `UVB index`
    )
  )
tbl_accessions_climate
```

```{r appendix = TRUE}
#| code-summary: "Export table tbl-accessions_climate"

# Export the table as a TSV file
readr::write_tsv(
  dta_accessions_climate_df,
  file = here::here("results", "tbl-accessions_climate.tsv"))

# Export the table as a LaTeX file
gt::gtsave(
  tbl_accessions_climate,
  filename = here::here("results", "tbl-accessions_climate.tex")
)
```

# Create maps of the bioclimatic data at the accession coordinates

@fig-accessions_map shows a map with the accessions in Spain and Sweden. @fig-accessions_annual_mean_temp_map, @fig-accessions_mean_diurnal_range_map, @fig-accessions_isothermality_map, @fig-accessions_temp_seasonality_map, @fig-accessions_temp_annual_range_map, @fig-accessions_annual_precipitation_map, and @fig-accessions_precipitation_seasonality_map, @fig-accessions_precipitation_driest_quarter_map show maps of the bioclimatic data at the accession coordinates: the annual mean temperature, the mean diurnal range, the isothermality, the temperature seasonality, the temperature annual range, the annual precipitation, the precipitation seasonality, and the precipitation of the driest quarter.

```{r appendix = TRUE}
#| code-summary: "Read in geographic data"

world_data <- rnaturalearth::ne_countries(scale = "large", returnclass = "sf")
europe_data <- world_data[world_data$continent == "Europe", ]

# Create a tibble of the points in Europe
europe_points <- dplyr::bind_cols(
  europe_data,
  sf::st_coordinates(sf::st_centroid(sf::st_make_valid(europe_data)$geometry))
)
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the accessions in Spain and Sweden"
#| label: fig-accessions_map
#| fig-cap: "Accessions in Spain and Sweden"
ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the annual mean temperature"
#| label: fig-accessions_annual_mean_temp_map
#| fig-cap: "Accessions in Spain and Sweden: annual mean temperature"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_1)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(0, 20)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Annual mean temperature"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the mean diurnal range"
#| label: fig-accessions_mean_diurnal_range_map
#| fig-cap: "Accessions in Spain and Sweden: mean diurnal range"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_2)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(5, 15)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Mean diurnal range"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the isothermality"
#| label: fig-accessions_isothermality_map
#| fig-cap: "Accessions in Spain and Sweden: isothermality"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_3)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(25, 45)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Isothermality"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the temperature seasonality"
#| label: fig-accessions_temp_seasonality_map
#| fig-cap: "Accessions in Spain and Sweden: temperature seasonality"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_4)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(450, 850)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Temperature seasonality"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the temperature annual range"
#| label: fig-accessions_temp_annual_range_map
#| fig-cap: "Accessions in Spain and Sweden: temperature annual range"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_7)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(20, 35)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Temperature annual range"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the annual precipitation"
#| label: fig-accessions_annual_precipitation_map
#| fig-cap: "Accessions in Spain and Sweden: annual precipitation"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_12)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(300, 1100)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Annual precipitation"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the precipitation seasonality"
#| label: fig-accessions_precipitation_seasonality_map
#| fig-cap: "Accessions in Spain and Sweden: precipitation seasonality"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_15)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(0, 100)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Precipitation seasonality"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

```{r appendix = TRUE}
#| code-summary: "Create a map of the precipitation of the driest quarter"
#| label: fig-accessions_precipitation_driest_quarter_map
#| fig-cap: "Accessions in Spain and Sweden:
#| precipitation of the driest quarter"

ggplot2::ggplot(data = europe_data) +
  ggplot2::geom_sf() +
  tidyterra::geom_spatraster(
    data = dta_bio_climate,
    mapping = ggplot2::aes(fill = wc2.1_30s_bio_17)
  ) +
  ggplot2::coord_sf(
    xlim = c(-10, 25),
    ylim = c(35, 70)
  ) +
  ggplot2::scale_fill_viridis_c(limits = c(0, 300)) +
  ggplot2::geom_point(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2
  ) +
  ggrepel::geom_text_repel(
    data = dta_accessions,
    mapping = ggplot2::aes(x = lon, y = lat, label = sample),
    color = ifelse(dta_accessions$pop == "ESP", "red", "darkblue"),
    size = 2.5,
    fontface = "bold",
    max.overlaps = Inf,
    min.segment.length = 0
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Spain"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = -2.8,
    nudge_y = -2.2
  ) +
  ggrepel::geom_text_repel(
    data = europe_points |> dplyr::filter(sovereignt == "Sweden"),
    mapping = ggplot2::aes(x = X, y = Y, label = name),
    fontface = "bold",
    min.segment.length = Inf,
    nudge_x = 2.8,
    nudge_y = 4.5
  ) +
  ggspatial::annotation_scale() +
  ggplot2::theme(
    panel.grid.major = ggplot2::element_line(
      color = gray(0.5),
      linetype = "dashed",
      linewidth = 0.5
    ),
    panel.background = element_rect(fill = "aliceblue")
  ) +
  ggplot2::labs(
    x = "Longitude",
    y = "Latitude",
    title = "Accessions in Spain and Sweden",
    subtitle = "Precipitation of the driest quarter"
  ) +
  ggplot2::theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 20),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14)
  )
```

# References {.unnumbered}

::: {#refs}
:::

# Appendix: Version information {.unnumbered}

::: {#r_pkg_version_info .callout-note collapse="true" appearance="minimal"}
## R session info with information on the R version as well as the packages that were loaded

```{r}
#| code-fold: false
#| message: true

message(paste(capture.output(sessionInfo()), "\n", collapse = ""))
```
:::

# Appendix: Source code {.unnumbered}

::: {#r_source_code .callout-note collapse="true" appearance="minimal"}
## The R source code for this research report

```{r ref.label=knitr::all_labels(appendix == TRUE)}
#| code-fold: false
#| eval: false
```
:::
