\documentclass[
11pt,
fleqn,
titlepage]{article}

% Packages

\usepackage[sfdefault]{roboto}
\usepackage{authblk}
\usepackage[a4paper, headheight=14pt]{geometry}
\usepackage[document]{ragged2e}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage[section]{placeins}
\usepackage{siunitx}
\usepackage{textgreek}
\usepackage[version=4]{mhchem}
\usepackage[babel, tracking, verbose]{microtype}
\usepackage[defaultlines=2, all]{nowidow}
\usepackage{subcaption}
\captionsetup{
    labelfont=bf,
    singlelinecheck=false,
    justification=RaggedRight}
\captionsetup[subfigure]{
    labelformat=brace}
\renewcommand\thesubfigure{\roman{subfigure}}
\usepackage{polyglossia}
\setdefaultlanguage[variant=american]{english}
\usepackage{csquotes}
\usepackage[style=apa]{biblatex}
\addbibresource{bibliography.bib}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{soul}
\usepackage[labelfont=bf]{caption}
\usepackage{tabularx}
\usepackage{threeparttablex}
\usepackage[imakeidx]{xindex}
\makeindex[intoc]
\usepackage[pdfa]{hyperref}
\usepackage{colorprofiles}
\hypersetup{
    pdfinfo={
        Title={An analysis of variations in drought stress-related genes of Arabidopsis thaliana - a comparison of populations from Spain and Sweden},
        Author={Halide Aydin, Sezin Dogan, Aliz Fodor, Janine Graser, Paula Winnitzki, Ingo Giebel},
        Subject={Research project QBio305: Population and Quantitative Genetics},
        Keywords={Heinrich-Heine-Universität Düsseldorf, Universität zu Köln, WS 2023/2024, QBio305: Population and Quantitative Genetics, Research project, Arabidopsis thaliana, Drought stress-related genes, Instructors: Prof. Dr. Juliette de Meaux, Dr. Markus Stetter, Dr. Tahir Ali}},
        colorlinks,
        linkcolor=blue,
        linktoc=all}
\usepackage[ocgcolorlinks, tikz]{ocgx2}
\usepackage{biocon}
\newplant{At}{genus=Arabidopsis, epithet=thaliana}
\sloppy

% Header

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\plant{At}: variations of drought stress-related genes in Spain and Sweden}
\fancyhead[R]{\thepage}
\fancypagestyle{plain}{}

% Title & author

\title{An analysis of variations in drought stress-related genes of \plant{At} - a comparison of populations from Spain and Sweden}
\author[1]{Halide Aydin, Sezin Dogan, Aliz Fodor, Janine Graser, Paula Winnitzki, Ingo Giebel}
\affil[1,2]{QBio305: Population and Quantitative Genetics}
\affil[1]{Heinrich-Heine-Universität Düsseldorf}
\affil[2]{Institute for Plant Sciences}
\affil[2]{Universität zu Köln}
\affil[2]{Prof. Dr. Juliette de Meaux, Dr. Markus Stetter, Dr. Tahir Ali}
\date{February 5, 2024}

\begin{document}
    \maketitle

    \tableofcontents

    \clearpage
    \renewcommand*\listtablename{List of tables}
    \listoftables
    \renewcommand*\listfigurename{List of figures}
    \listoffigures

    \clearpage

    \input{abstract}

    \input{introduction}

    \input{materials_and_methods}

    \input{results}

    \input{discussion}

    \clearpage
    \printbibliography[heading=bibintoc]

    \clearpage
    \appendix
    \input{appendix_a}
    \input{appendix_b}
    \input{appendix_c}

    \clearpage
    \printindex
\end{document}